---
title: "Concept rule learning with an ontology-based refinement operator
inducing semnatic rules from omics data"
author: "Frantisek Malinka"
output:
BiocStyle::html_document:
toc_float: true
vignette: >
%\VignetteIndexEntry{sem1R - A package inducing semantic rules from omics data}
%\VignetteEngine{knitr::rmarkdown}
%\VignetteEncoding{UTF-8}
%\VignetteDepends{sem1R, BiocStyle}
%\VignettePackage{sem1R}
%\VignetteKeywords{semantic, rule learning, clustering, enrichment}

---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

```{r init, message = FALSE, echo = FALSE, results = "hide" }
## Silently loading all packages
library(BiocStyle)
library(sem1R)

```

**Package**: `r Biocpkg("sem1R")`<br />
**Authors**: Frantisek Malinka<br />
**Modified**: `r file.info("sem1R.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

sem1R is a machine learning algorithm that finds interesting, hidden,
and non-trivial patterns in omics data. The algorithm produces
a set of semantical prediction rules that form data into clusters
or biclusters, this depends on a type of ontologies. Here, we distingues
between two types of ontologies: an ontology describing rows (e.g. genes)
an columns (e.g. samples). Practically, for gene expression data, where
rows represent genes and column represent samples, we recommend to use
Gene ontology or any pathway ontologies as a row ontology. Choosing
a proper column ontology is depending on a type of experiment,
e.g. OBO Foundry provides almost two hundreds and many of them are domain
specific so some anatomical ontologies can be used as well.

The sem1R is based on rule learning methods, where two reduction
procedures were added and that make the algorithm extremely fast
and efficient in comparison with traditional approach. In additional,
it is relative easy to use, because all important methods are included
into the package.

# Getting Started

The algorithm is implemented in C++ and provided as R package.
The following instructions will show you how to install all prerequisites
and the sem1R package as well into your local machine. Afterwards, we will
demonstrate the sem1R on real gene expression dataset.

# Prerequisites

We required to use R in version 3.0.2. All prerequisites R packages
that are needed for the sem1R package are the following: Rcpp (>= 0.12.6),
RcppProgress, RcppArmadillo (>= 0.7.800.2.0), and BH (>= 1.72.0-3).
All of these packages come from CRAN, so install them by install.packages
function in R.

Or, for easier installation we recommend to install 'devtools' that can
download and install the project instantly from gitHub using only one command.

For installing 'devtools' package run R and type the following:

install.packages("devtools")

# Installing

If you choosen instalation via devtools, you would go to the terminal
and run R and then the following commands:

>library(devtools)
>install_gitHub("fmalinka/sem1R")

All prerequisites packages should be installed automatically.

For non-devtools users, download the sem1R package and extract it to
your arbitrary folder. Run the terminal, go to the folder that you have
already choosen and install all prerequisities if you have not installed
yet using install.packages R's function. Go to the folder and build
the package.

cd /my/path/to/package
R CMD build .

The package in tar.gz format will be named as `sem1R_[version].tar.gz'.
The concrete name depends on the package version. Then, install
the sem1R package.

R CMD INSTALL sem1R_[version].tar.gz

And finally, check whether the sem1R package has been installed.
Run R and load the package.

>library(sem1R)

